---
title: "knmitransformer - examples"
author: "Martin Roth & Christiana Photiadou"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{StandardStations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r LoadLibraries, include=FALSE}
library(data.table)
library(knmitransformer)
```
The knmitransformer pacakge can be used with both *standard stations*, as the ones used in the published Climate Scenarions by KNMI, and *user definied stations*. This document provides examples for both situations and gives step by step examples for calculating transformed series of precipitation, temperature, radiation and Makkink evaporation.

## Standard Stations
Standard Stations refer to the stations already used in the published climate scenarios with reference period 1981-2010. These reference data can be also found in the [KNMI'14-klimaatscenario's webpage](http://www.klimaatscenarios.nl/toekomstig_weer/transformatie/index.html)

 
**Precipitation**

To calculate transformed precipitation series using the reference data from the period 1981-2010, we have to specify the scenario, horizon, and subscenario (plese use ``` ?TransformPrecip ``` for more information) 
```{r precipSelection}
subscenario <- "centr"
input       <- system.file("refdata",
                           "KNMI14____ref_rrcentr___19810101-20101231_v3.2.txt",
                           package="knmitransformer")
ofile       <- "tmp.txt" # output file - used only temporary
scenario    <- "GL"
horizon     <- 2030
```

```{r precipTransformation}
futPrecip <- TransformPrecip(input=input, ofile=NA,
                             scenario=scenario, horizon=horizon,
                             subscenario=subscenario)
```

```{r precipShowResult}
head(futPrecip[, 1 : 5])
```

The first five lines of columns V2 - V5 of the transformed input represent the
station number, rd northing and rd easting, lat and lon of each station. 

For Temperature, Radiation and Makkink evaporation similar steps are needed:

**Temperature**
```{r tempSelection}
var      <- "tg" #for daily average temperature; other options are daily maximum (tx), minimum (tn) temperature
input    <- system.file("refdata",
                        "KNMI14____ref_tg___19810101-20101231_v3.2.txt",
                        package="knmitransformer")
regions  <-  MatchRegionsOnStationId(ReadInput(var, input)$header[1, -1])
scenario <- "GL"
horizon  <- 2030
```

```{r tempTransformation}
futTemp <- TransformTemp(input=input, ofile=NA,
                         scenario=scenario, horizon=horizon,
                         var=var, regions=regions)
```

```{r tempShowResult}
head(futTemp[, 1 : 5])
```

**Radiation**
```{r radSelection}
input      <- system.file("refdata",
                        "KNMI14____ref_rsds___19810101-20101231_v3.2.txt",
                        package="knmitransformer")
scenario   <- "GL"
horizon    <- 2030
```

```{r radTransformation}
futRad <- TransformRadiation(input=input, ofile=NA,
                         scenario=scenario, horizon=horizon)
```

```{r radShowResult}
head(futRad[, 1 : 5])
```

**Makkink evaporation**
```{r evapSelection}
inputTemp   <- system.file("refdata",
                          "KNMI14____ref_tg___19810101-20101231_v3.2.txt",
                          package="knmitransformer")
inputRad <- system.file("refdata",
                        "KNMI14____ref_rsds___19810101-20101231_v3.2.txt",
                        package="knmitransformer")
scenario   <- "GL"
horizon    <- 2030
regions  <-  MatchRegionsOnStationId(ReadInput("tg", inputTemp)$header[1, -1])
```

```{r evapTransformation}
futEvap <- TransformEvap(inputTemp=inputTemp, inputRad=inputRad,
                         scenario=scenario, horizon=horizon,
                         regions = regions)
```

```{r evapShowResult}
head(futEvap[, 1 : 5])
```

## User definied stations

User definied stations, such as the data set in this example for daily temperature data set, which was downloaded from the [Climate Explorer](https://climexp.knmi.nl/selectdailyseries.cgi?id=someone@somewhere) will need some actions beforehand. In this example data sets from De Bilt station are used.

As you can see the data set has a slightly different format:
```{r LoadTGClimateExplorer, echo = FALSE}
filename <- system.file("exampledata",
                         "tgtg260.dat",
                         package="knmitransformer")
tgData <- fread(filename, skip = 6)
head(tgData, n = 10)
```

For such data sets we will need to define an extra date column, provide latitutde and longitude details, and select an appropriate base period of 30 years. A 30 year period is important for statistical stability; keep in mind that the transformation programme was written with the base period 1981 - 2010 in mind.

First, we append one single date column, which is necessary:
```{r CorrectDate}
tgData[, date := as.integer(format(as.Date(paste(V1, V2, V3, sep = "-"),
                        format = "%Y-%m-%d"), format = "%Y%m%d"))]
head(tgData, n = 10)
```
Then we subset the correct period and ensure that the column names are date and values. Latitude and longitude are provided as well.

```{r FinalPreparations}
tgData <- tgData[V1 %in% 1981 : 2010, .(date, values = V4)]
head(tgData, n = 10)
lat <- 52.1
lon <- 5.18
```

Now we can create the input object and transform it in the next step (use ?CreatKnmiTFInput for more information)
```{r CreateValidatedInput}
input <- CreateKnmiTFInput(tgData, 1, lat, lon)
inputTrans <- TransformTemp(input, scenario = "GL", var = "tg", horizon = 2030,
                            regions = "MON") 
head(inputTrans, n = 15)
```

The first five lines of the second column of the transformed input represent the
station number, rd northing and rd easting, lat and lon. 

For precipitation we can perform a similar exercise ...
```{r}
filename <- system.file("exampledata",
                        "hom1951550.dat",
                        package="knmitransformer")
precipData <- fread(filename, skip = 6)

setnames(precipData, c("date", "values"))

precipData <- precipData[date %/% 10000 %in% 1981 : 2010, ]
head(precipData, n = 10)

input <- CreateKnmiTFInput(precipData, 550, lat, lon)

inputTrans <- TransformPrecip(input, scenario = "GL", horizon = 2030)
head(inputTrans, n = 15)
```

And for radiation also...
```{r}
filename <- system.file("exampledata",
                        "qqqq260.dat",
                        package="knmitransformer")
radiationData <- fread(filename, skip = 6)

head(radiationData, n = 10)

radiationData[, date := as.integer(format(as.Date(paste(V1, V2, V3, sep = "-"),
                        format = "%Y-%m-%d"), format = "%Y%m%d"))]
head(radiationData, n = 10)

radiationData <- radiationData[V1 %in% 1981 : 2010, .(date, values = V4)]

head(radiationData, n = 10)
```
The unit here is w/m2 instead of kJ/m2. Therefore, we have to adapt the data
first
```{r}
radiationData[, values := values * 86.4]
head(radiationData, n = 10)
```

```{r}

input <- CreateKnmiTFInput(radiationData, 260, lat, lon)

inputTrans <- TransformRadiation(input, scenario = "GL", horizon = 2030) 
head(inputTrans, n = 15)

```


  